---
title: "plot2() plotting"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{plot2() plotting}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(certeplot2)
library(ggplot2)
library(dplyr)
```

# Setting plot type / geom

# Setting the 3 directions 

## Setting multiple columns {.tabset}

Support multiple columns, with `c(...)` or even Tidyverse selectors:

### `plot2`

```{r}
admitted_patients %>%
  plot2(x = hospital, facet = where(is.character))
```

### `ggplot2`

```{r}
admitted_patients %>%
  ggplot(aes(x = hospital, y = age)) +
  geom_boxplot() +
  # the facets require to call the data again
  # (or requires extensive mutate()/across() code)
  facet_wrap(facets = admitted_patients %>%
               select(where(is.character)) %>%
               colnames())
```

## The x axis

## The category

## The facet

# Setting the values

# Sorting {.sorting}

# Transforming

# Legends

# Plotting geometries (`sf` objects)

Objects of class `sf` (*simple feature*) are treated specially by the `ggplot2` package, although `plot2()` handles these objects like any other data set. Instead it returns a `ggplot` object created with `geom_sf()` internally, but most arguments are usable and even the `theme_minimal2()` theme has good use for `sf` plots.

The `netherlands` data set is such an `sf` object:

```{r}
library(sf)
print(netherlands)
```

How does it look?

```{r}
plot2(netherlands)
```
